Quasi-Monte Carlo Methods: What, Why, and How?

### 2025-01-09 HickernellMCQMC2024_QMCTutorial.tex
M2 Mac updates

%%Fred's temporary macros
\newcommand{\figpath}{Figures}

@@

Let's start with the one-dimensional extensible lattice that is defined by the van der Corput sequence, $\{\phi_b(0), \phi_b(1), \ldots\}$.  This involves reflecting the digits of the integer $i$ in base $b$ about the decimal point.  For example, $i = 6 = 110_{2}$ gives $\phi_2(6) = {}_20.011 = 3/8$, or in general,
\begin{equation}
	i = i_0 + i_1b + i_2 b^2 + \cdots \implies \phi_b(i) := i_0 b^{-1} + i_1 b^{-2} + i_2 b^{-3} + \cdots \in [0,1).
\end{equation}
The first $n = b^m$ nodes in the van der Corput sequence correspond to the evenly spaced nodes $\{0, b^{-m}, \ldots, 1 - b^{-m} \}$---albeit in a different order---for all positive integers $m$.

@@

For the example in Figure \ref{fig:extensiblelatticeconstruct}, increasing $n$ beyond $16$ leads to repetition of the orginal points because the generator $\bsh$ has integers less than $16$.  Obtaining a truly extensible lattice sequence that  the

@@

>>>>

%%Fred's temporary macros
\newcommand{\figpath}{Figures}
\usepackage{xspace}

@@

Let's start with the one-dimensional extensible lattice that is defined by the van der Corput sequence, $\{\phi_b(0), \phi_b(1), \ldots\}$.  This involves reflecting the digits of the integer $i$ in base $b$ about the decimal point.  For example, $i = 6 = 110_{2}$ gives $\phi_2(6) = {}_20.011 = 3/8$, or in general,
\begin{equation} \label{eq:vdc}
	\phi_b(i) := i_0 b^{-1} + i_1 b^{-2} + i_2 b^{-3} + \cdots \in [0,1) \quad \text{for }
		i = i_0 + i_1b + i_2 b^2 + \cdots .
\end{equation}
For all non-negative integers, $m$, the first $n = b^m$ nodes in the van der Corput sequence correspond to the evenly spaced nodes $\{0, b^{-m}, \ldots, 1 - b^{-m} \}$---albeit in a different order.

@@

@@

For the example in Figure \ref{fig:extensiblelatticeconstruct}, increasing $n$ beyond $16$ repeats the orginal points because the generating vector $\bsh$ has integers less than $16$.  Obtaining a truly extensible lattice sequence requires that $\bsh$ be a vector of generalized integers---essentially integers with infinite numbers of nonzero digits---as explained in \cite{HicNie03a}, where the existence of good generating vectors is also proved.  In practice, one searches computationally for a $\bsh$ that produces good set of lattice nodes of size $n = b^m$ for a range of $m$ \cite{HicEtal00}.

\subsection{Digital Sequences} \label{sec:digital}

Another family of LD sequences that also builds from the van der Corput sequence of \eqref{eq:vdc} is digital sequences \cite{DicPil10a,Nie92}.  For simplicity, we restrict ourselves to $b = 2$ and let $\oplus$ denote digitwise addition, e.g., $3/8 \oplus 3/4 = {}_20.011 \oplus {}_20.110 = {}_20.101 = 5/8$.  A digital sequence is defined as
\begin{equation} \label{eq:vdc}
	\bsx_i := i_0 \bsc_1 \oplus i_1 \bsc_2 \oplus i_2 \bsc_3 \oplus \cdots \in [0,1)^d \quad \text{for }
	i = i_0 + i_12 + i_2 2^2 + \cdots,
\end{equation}
where $\bsc_1, \bsc_2, \ldots$ are carefully chosen.  If $c_{ij\ell}$ denotes the $\ell^{\text{th}}$ digit of the $j^{\text{th}}$ coordinate of $\bsc_i$, then in the literature one may see the generators of digital sequences stored as generating matrices, $\mathsf{C}_j = (c_{ij\ell})_{\ell,i = 1}^{M,N}$ for $j = 1, \ldots, d$, where $M$ is th maximum number of bits in the expression for $\bsx_i$, say $52$, and $2^N$ is the maximum number of nodes that is intended to be generated.

Digital sequence generators can be found via number theory or numerical search.  The
